---
title: Extending Department Capability With GIS
author: Martin Gleason, MS
output:
  xaringan::moon_reader:
    css: ["libs/remark-css/default.css", "pro_presentation.css"]
    self_contained: no
    lib_dir: libs
    seal: false
    nature:
      beforeInit: "macros.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include= FALSE, echo = FALSE, warning=FALSE}
library(kableExtra)
source("cleaning_aoic.R")
source("gis_demo.R")
source("cta_data.R")
```
# The Ask

  * Cook County Geographic Information Systems (CCGIS) department is willing to develop a number of dashboard and applications for the department 

* They require some of the department's data to build their solutions
  * The data has no externally identifying information
  * The data would be housed on their servers during protoytping
  
* Working with CCGIS will enhance the value of CFive and improve the department's analytical capability

* Cost factors have already been budgeted for this in FYI 2019
  * These are all operational expenses, as no new software or hardware needs to be purchased

---
# GIS

* GIS stands for Geographic Information Services
  * [Cook Central](https://hub-cookcountyil.opendata.arcgis.com)
  
* Cook County has an deparment that has worked with:
  * [Cook County Sheriff's department](https://maps.cookcountyil.gov/sheriffreentryassistance/)
  * [Cook County Medical Examiner's Office](https://maps.cookcountyil.gov/medexammaps/)
  * [Cook County Assessors](http://hub-cookcountyil.opendata.arcgis.com/datasets/5ec856ded93e4f85b3f6e1bc027a2472_0)

* The current director of CCGIS is a long time supporter of [MapCORPS](https://mapscorps.org)

---
# CCGIS and the Department

.pull-left[
* CCGIS needs sample data that has:
  * Addresses
  * Charges
  * Supervision Level
  * Final Status
  
* No identifying information will be given to CCGIS
  * Data can be a file and not server access
  * Server access should be discussed if prototypes meet department needs
]

.pull-right[
* CCGIS can create applications with data that:
  * Assist in automatically assigning cases based on geographic location
  * Shows gaps in services and dosage throught the county
  * Clusters of needs
  * Systemtemic barriers to services
      * Average time to service providers
      * Client to service provider ratio
]

---

## GIS and the Department: Demo
  
* The following demonstrations were developed with sample data used in the Discretion Mapping Project
  * Data was stripped of all identifying information
    * The data was not cleaned<sup>1</sup>
  * Addresses were geocoded by Google Maps API
  
* Mock-ups were created with:
  * [R](https://www.r-project.org), [Leaflet JS](https://leafletjs.com), and [Google's API](https://developers.google.com/maps/documentation/)
  * CCGIS uses to [ARCGis](https://www.arcgis.com/index.html), a complete GIS solution
  * All of the tools we would use are either
     *  Free
     *  Already under contract
     *  Maps would be interactive with proper judicial permission

.footnote[1: Clean data removes typos and addressses missing values]

---
## GIS and the Department
.left-plot[
```{r gis_chicago_pd, include = TRUE, echo = FALSE, warning = FALSE, out.width="100%", fig.height =6, fig.align='left'}

chicago_with_districts_ggmap

```
]

.right-code[
* A map of Chicago Metro Area 
  * With the Chicago Police Department Districts 
  * Police district data comes from CPD
  * Chaging boundaries would be a simple task
  * CCGIS would use muncipal boundaries for suburban districts
]

---
## Diversions: From `r min(geo_coded_diversions$STRDTE)` to `r max(geo_coded_diversions$ENDDTE, na.rm = TRUE)`

```{r gis_demo, include = TRUE, echo = FALSE, out.width="95%", fig.height =6}
all_diversions

```

---
# Understanding the Map

* All `r nrow(geo_coded_diversions_sf)` Diversions
  *  From 2012 to 2017
  *  The year 9999 is clearly a placeholder value

* Indicates where diversions occured *throughout* the country

* Currently, the data is limited to: `r names(geo_coded_diversions[1:7])`
* ITVSTS appears to be "Intervention Status"

* Filter for the most recent cases gives us a more useable map
* A total of `r nrow(filter(geo_coded_diversions_sf, lubridate::year(STRDTE) >= 2016))` clients

---
## Intervention Status: 2016-2017

```{r intervention_status, include=TRUE, echo = FALSE, warning=FALSE, out.width="100%", fig.height =6}

leaflet(cpd_districts) %>%
  addTiles() %>%
  setView(lng = -87.7, lat = 41.9, zoom = 12) %>%
  addPolygons(color = "black",
              fillColor = "#D3D3D3", 
              fillOpacity = 0.1) %>%
  addCircles(data = filter(geo_coded_diversions_sf, lubridate::year(STRDTE) >= 2016),
             color = ~pal(ITVSTS)) %>%
  addLegend("bottomright", pal = pal,
            value= ~geo_coded_diversions_sf$ITVSTS,
            title = "Intervention Status") 
```

# Chicago: Diversions with CTA Stops