---
title: Extending Department Capability With GIS
author: Martin Gleason, MS
output:
  xaringan::moon_reader:
    css: ["libs/remark-css/default.css", "pro_presentation.css"]
    self_contained: no
    lib_dir: libs
    seal: false
    nature:
      beforeInit: "macros.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include= FALSE, echo = FALSE, warning=FALSE}
library(kableExtra)
source("cleaning_aoic.R")
source("cta_data.R")
source("gis_demo.R")

```

.center[# Data Sharing With CCGIS]
* Cook County Geographic Information Systems (CCGIS) department is willing to develop a number of dashboard and applications for the department 

* They require some of the department's data to build their solutions

* Working with CCGIS will enhance the value of CFive and improve the department's analytical capability

---
# Data Sharing Concerns

* Cook County Juvenile Court data has specific privacy restrictions
  * Proposed data for prototyping has no externally identifying information
  * The data would be housed on their servers during prototyping
    * Data product would be hosted internal

* CCGIS needs sample data that has:
  * Addresses
  * Charges
  * Supervision Level
  * Final Probation/Supervision Status
  
---
# Cook County GIS

* GIS stands for Geographic Information Services
  * [Cook Central](https://hub-cookcountyil.opendata.arcgis.com)
  
* Cook County has an department that has worked with:
  * [Cook County Sheriff's department](https://maps.cookcountyil.gov/sheriffreentryassistance/)
  * [Cook County Medical Examiner's Office](https://maps.cookcountyil.gov/medexammaps/)
  * [Cook County Assessors](http://hub-cookcountyil.opendata.arcgis.com/datasets/5ec856ded93e4f85b3f6e1bc027a2472_0)

* The current director of CCGIS is a long time supporter of [MapCORPS](https://mapscorps.org)

---
# CCGIS and the Department

* CCGIS can create applications with data that:
  * Assist in automatically assigning cases based on geographic location
  * Shows gaps in services and dosage throughout the county
  * Clusters of needs
  * Systemic barriers to services
      * Average time to service providers
      * Client to service provider ratio

---
# GIS and the Department: Demo
  
* The following demonstrations were developed with sample data used in the Discretion Mapping Project
  * Data was stripped of all identifying information
    * The data was not cleaned<sup>1</sup>
  * Addresses were geocoded by Google Maps API
  
* Mock-ups were created with:
  * [R](https://www.r-project.org), [Leaflet JS](https://leafletjs.com), and [Google's API](https://developers.google.com/maps/documentation/)
  * CCGIS uses to [ARCGis](https://www.arcgis.com/index.html), a complete GIS solution
  * All of the tools we would use are either
     *  Free
     *  Already under contract
     *  Interactive only with proper judicial permission

.footnote[1: Clean data removes typos and addresses missing values]

---
# GIS and the Department
.left-plot[
```{r gis_chicago_pd, include = TRUE, echo = FALSE, warning = FALSE, out.width="100%", fig.height =6, fig.align='left'}

chicago_with_districts_ggmap +
  labs(caption = "Source: https://data.cityofchicago.org/Public-Safety/CPD-districts/7hhi-ktqw")

```
]

.right-code[
* Police district data comes from CPD<sup>1</sup>

  *  Changing boundaries would be a simple task

* CCGIS would use municipal boundaries for suburban districts
]

---
##Diversions<sup>1</sup> from `r as.Date(geo_coded_diversions$STRDTE) %>% min() %>% format("%m-%d-%Y")` to `r max(geo_coded_diversions$ENDDTE, na.rm = TRUE) %>% format("%m-%d-%Y")`

```{r gis_demo, include = TRUE, echo = FALSE, out.width="95%", fig.height =6}
all_diversions

```

---
# Understanding the Map

* All `r nrow(geo_coded_diversions_sf)` Diversions
  *  From 2012 to 2017
  *  The year 9999 is clearly a placeholder value

* Indicates where diversions occurred *throughout* the country
  * And one person from England

* Currently, the data is limited to: `r names(geo_coded_diversions[1:7])`

* ITVSTS appears to be "Intervention Status"


.footnote[1: Data from JEMS for the Discretion Mapping Project]
---
## Intervention Status: 2016-2017

```{r intervention_status, include=TRUE, echo = FALSE, warning=FALSE, out.width="100%", fig.height =6}

leaflet(cpd_districts) %>%
  addTiles() %>%
  setView(lng = -87.7, lat = 41.9, zoom = 12) %>%
  addPolygons(color = "black",
              fillColor = "#D3D3D3", 
              fillOpacity = 0.1) %>%
  addCircles(data = filter(geo_coded_diversions_sf, lubridate::year(STRDTE) >= 2016),
             color = ~pal(ITVSTS)) %>%
  addLegend("bottomright", pal = pal,
            value= ~geo_coded_diversions_sf$ITVSTS,
            title = "Intervention Status from JEMS") 
```
---
###  Intervention Status Map: Decoded
```{r caluclations, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

n_clients <- geo_coded_diversions_sf%>%
  filter(lubridate::year(STRDTE) >= 2016) %>%
  nrow()
```
* Filtering for the most recent cases gives us a more usable map

* A total of `r n_clients` clients

* Partnering with GIS would allow for more real time analytics:
  *  Based on current JEMS/CFive Data
  *  And data from other sources
    *  CTA
    *  DCFS
    *  Community partners

---
# Additional Sources Example
.right-plot[
```{r heatmap, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, out.width="100%", fig.height =6, fig.align='right'}
heat_map_CTA +
  labs(caption = "Data source is from previously listed CTA and CPD sources")
```
]
.left-code[
 
* CTA Stop Data

* Police District Data

* Calculated in R 

* More visualizations like this could occur with additional collaboration with CCGIS

* This all begins when we share data with our sister agency
]
